# Engineering professionals

Average monthly pay, non-manual workers private sector (SLP) by region, occupational group (SSYK) and sex. Year 2000 - 2013  
Average monthly pay (total pay), non-manual workers private sector (SLP), SEK  
214 Engineering professionals	  
men and women	  
      
```{r Engineeringpro00-13, echo = TRUE, fig.cap = 'Engineering professionals, Year 2000 - 2013'}  
readfile ("AM0103H2.csv") %>%   
  ggplot () +  
    geom_line (mapping = aes (x = year2, y = salary, colour = region)) +  
    theme (legend.position = "bottom")  

readfile ("AM0103H2.csv") %>%   
  group_by (region) %>%
  mutate (growth = c (NA, diff(salary)) / salary) %>%  
  ggplot () +    
    geom_line (mapping = aes(x = year2, y = growth, colour = region)) +  
    theme (legend.position = "bottom")	  
```

Average monthly pay, non-manual workers private sector (SLP) by region, occupational group (SSYK 2012) and sex. Year 2014 - 2018  
Average monthly pay (total pay), non-manual workers private sector (SLP), SEK  
214 Engineering professionals	  
men and women  
	
```{r Engineeringpro14-18, echo = TRUE, fig.cap = 'Engineering professionals, Year 2014 - 2018'}  	
readfile ("0000002T.csv") %>%   
  ggplot () +  
    geom_line (mapping = aes (x = year2, y = salary, colour = region)) +  
    theme (legend.position = "bottom")  
	
readfile ("0000002T.csv") %>%   
  group_by (region) %>%
  mutate (growth = c (NA, diff(salary)) / salary) %>%  
  ggplot () +    
    geom_line (mapping = aes(x = year2, y = growth, colour = region)) +  
    theme (legend.position = "bottom")	
```