# R&D expenditures

## 

Employees 16-64 years by region of work, occupation (3-digit SSYK 2012) and sex. Year 2014 - 2017
214 Engineering professionals

Population 16-74 years of age by region, highest level of education, age and sex. Year 1985 - 2018
total, 16-74 years

Regional distribution of enterprises R&D. Every other year 2007 - 2017
Intramural R&D, SEK million

Total R&D person years performed by employees of a higher education institution by region, sex and field of science. Every second year 2013 - 2017
Total R&D person years
County
men, women
All fields of science and technology


```{r rndexpenditures1, echo = TRUE, fig.cap = 'correlation matrix between R&D expenditures and number of engineers and level of education'} 
tb <- readfile("000000NL_1.csv") %>%
  group_by (`region`, year) %>%  
  mutate (perc_women = as.numeric (sub ("%", "", perc_women (salary)))) %>%
  mutate (sum_ing = sum(salary))  
tb <- readfile("UF0506A1_2.csv") %>%
  group_by(`level of education`, region, year, sex) %>%
  mutate(utbregno = sum(salary)) %>%
  group_by(region, year, sex) %>% mutate(perc_edu = utbregno / sum(utbregno)) %>%
  group_by(`level of education`, region, year) %>%
  mutate (sum_edu = sum(utbregno)) %>%
  filter (`level of education` == "post-secondary education 3 years or more (ISCED97 5A)") %>%
  right_join(tb, by = c("region", "year", "sex")) %>%
  mutate (perc_eng = sum_ing / sum_edu)
tb <- readfile("000002O5.csv") %>%
  right_join(tb, by = c("region", "year"))
tb <- readfile("000002LR.csv") %>%
  right_join(tb, by = c("region", "year"))
  
tb1 <- tb %>% select(
  salary.x.x, 
  salary.y, 
  salary.x, 
  year2.x, 
  utbregno, 
  perc_edu, 
  sum_edu, 
  salary.y.y, 
  perc_women, 
  sum_ing, 
  perc_eng)

chart.Correlation(tb1, histogram=TRUE, pch=19)
```

```{r rndexpenditures2, echo = TRUE, fig.cap = 'clustering tendancy for data'} 
tb <- readfile("000000NL_1.csv") %>%
  group_by (`region`, year) %>%  
  mutate (perc_women = as.numeric (sub ("%", "", perc_women (salary)))) %>%
  mutate (sum_ing = sum(salary))  
tb <- readfile("UF0506A1_2.csv") %>%
  group_by(`level of education`, region, year, sex) %>%
  mutate(utbregno = sum(salary)) %>%
  group_by(region, year, sex) %>% mutate(perc_edu = utbregno / sum(utbregno)) %>%
  group_by(`level of education`, region, year) %>%
  mutate (sum_edu = sum(utbregno)) %>%
  filter (`level of education` == "post-secondary education 3 years or more (ISCED97 5A)") %>%
  right_join(tb, by = c("region", "year", "sex")) %>%
  mutate (perc_eng = sum_ing / sum_edu)
tb <- readfile("000002O5.csv") %>%
  right_join(tb, by = c("region", "year"))
tb <- readfile("000002LR.csv") %>%
  right_join(tb, by = c("region", "year"))
  
tb1 <- tb %>% select(
  salary.x.x, 
  salary.y, 
  salary.x, 
  year2.x, 
  utbregno, 
  perc_edu, 
  sum_edu, 
  salary.y.y, 
  perc_women, 
  sum_ing, 
  perc_eng)
  
get_clust_tendency(
  tb1, 
  n = 50,
  gradient = list(low = "steelblue",  high = "white"))  
```				   
				   
```{r rndexpenditures3, echo = TRUE}
tb <- readfile("000000NL_1.csv") %>%
  group_by (`region`, year) %>%  
  mutate (perc_women = as.numeric (sub ("%", "", perc_women (salary)))) %>%
  mutate (sum_ing = sum(salary))  
tb <- readfile("UF0506A1_2.csv") %>%
  group_by(`level of education`, region, year, sex) %>%
  mutate(utbregno = sum(salary)) %>%
  group_by(region, year, sex) %>% mutate(perc_edu = utbregno / sum(utbregno)) %>%
  group_by(`level of education`, region, year) %>%
  mutate (sum_edu = sum(utbregno)) %>%
  filter (`level of education` == "post-secondary education 3 years or more (ISCED97 5A)") %>%
  right_join(tb, by = c("region", "year", "sex")) %>%
  mutate (perc_eng = sum_ing / sum_edu)
tb <- readfile("000002O5.csv") %>%
  right_join(tb, by = c("region", "year"))
tb <- readfile("000002LR.csv") %>%
  right_join(tb, by = c("region", "year"))
  
tb1 <- tb %>% select(
  salary.x.x, 
  salary.y, 
  salary.x, 
  year2.x, 
  utbregno, 
  perc_edu, 
  sum_edu, 
  salary.y.y, 
  perc_women, 
  sum_ing, 
  perc_eng)
  
tb2 <- tb1 %>% na.omit() %>% scale()  
  
intern <- clValid(tb2, nClust = 2:6, 
              clMethods = c("hierarchical","kmeans","pam"),
              validation = "internal")
			  
summary(intern) %>% 
  tidy() %>% 
  knitr::kable( 
    booktabs = TRUE,
    caption = 'Clustering validation measures')			  
```

```{r rndexpenditures4, echo = TRUE, fig.cap = 'Hierarchical clustering using Ward's method'} 
tb <- readfile("000000NL_1.csv") %>%
  group_by (`region`, year) %>%  
  mutate (perc_women = as.numeric (sub ("%", "", perc_women (salary)))) %>%
  mutate (sum_ing = sum(salary))  
tb <- readfile("UF0506A1_2.csv") %>%
  group_by(`level of education`, region, year, sex) %>%
  mutate(utbregno = sum(salary)) %>%
  group_by(region, year, sex) %>% mutate(perc_edu = utbregno / sum(utbregno)) %>%
  group_by(`level of education`, region, year) %>%
  mutate (sum_edu = sum(utbregno)) %>%
  filter (`level of education` == "post-secondary education 3 years or more (ISCED97 5A)") %>%
  right_join(tb, by = c("region", "year", "sex")) %>%
  mutate (perc_eng = sum_ing / sum_edu)
tb <- readfile("000002O5.csv") %>%
  right_join(tb, by = c("region", "year"))
tb <- readfile("000002LR.csv") %>%
  right_join(tb, by = c("region", "year"))
  
tb1 <- tb %>% select(
  salary.x.x, 
  salary.y, 
  salary.x, 
  year2.x, 
  utbregno, 
  perc_edu, 
  sum_edu, 
  salary.y.y, 
  perc_women, 
  sum_ing, 
  perc_eng)
  
tb2 <- tb1 %>% na.omit() %>% scale()  
  
d <- dist(tb2, method = "euclidean")

res.hc <- hclust(d, method = "ward.D2" )

grp <- cutree(res.hc, k = 2)

plot(res.hc, cex = 0.6) # plot tree
rect.hclust(res.hc, k = 2, border = 2:5)
```

```{r rndexpenditures4, echo = TRUE, fig.cap = 'Hierarchical clustering using Ward's method'} 
tb <- readfile("000000NL_1.csv") %>%
  group_by (`region`, year) %>%  
  mutate (perc_women = as.numeric (sub ("%", "", perc_women (salary)))) %>%
  mutate (sum_ing = sum(salary))  
tb <- readfile("UF0506A1_2.csv") %>%
  group_by(`level of education`, region, year, sex) %>%
  mutate(utbregno = sum(salary)) %>%
  group_by(region, year, sex) %>% mutate(perc_edu = utbregno / sum(utbregno)) %>%
  group_by(`level of education`, region, year) %>%
  mutate (sum_edu = sum(utbregno)) %>%
  filter (`level of education` == "post-secondary education 3 years or more (ISCED97 5A)") %>%
  right_join(tb, by = c("region", "year", "sex")) %>%
  mutate (perc_eng = sum_ing / sum_edu)
tb <- readfile("000002O5.csv") %>%
  right_join(tb, by = c("region", "year"))
tb <- readfile("000002LR.csv") %>%
  right_join(tb, by = c("region", "year"))
  
tb1 <- tb %>% select(
  salary.x.x, 
  salary.y, 
  salary.x, 
  year2.x, 
  utbregno, 
  perc_edu, 
  sum_edu, 
  salary.y.y, 
  perc_women, 
  sum_ing, 
  perc_eng)
  
tb2 <- tb1 %>% scale()  
  
d <- dist(tb2, method = "euclidean")

res.hc <- hclust(d, method = "ward.D2" )

grp <- cutree(res.hc, k = 2)

plot(res.hc, cex = 0.6) # plot tree
rect.hclust(res.hc, k = 2, border = 2:5)
```

```{r rndexpenditures4, echo = TRUE, fig.cap = 'Hierarchical clustering using Ward's method'} 
tb <- readfile("000000NL_1.csv") %>%
  group_by (`region`, year) %>%  
  mutate (perc_women = as.numeric (sub ("%", "", perc_women (salary)))) %>%
  mutate (sum_ing = sum(salary))  
tb <- readfile("UF0506A1_2.csv") %>%
  group_by(`level of education`, region, year, sex) %>%
  mutate(utbregno = sum(salary)) %>%
  group_by(region, year, sex) %>% mutate(perc_edu = utbregno / sum(utbregno)) %>%
  group_by(`level of education`, region, year) %>%
  mutate (sum_edu = sum(utbregno)) %>%
  filter (`level of education` == "post-secondary education 3 years or more (ISCED97 5A)") %>%
  right_join(tb, by = c("region", "year", "sex")) %>%
  mutate (perc_eng = sum_ing / sum_edu)
tb <- readfile("000002O5.csv") %>%
  right_join(tb, by = c("region", "year"))
tb <- readfile("000002LR.csv") %>%
  right_join(tb, by = c("region", "year"))
  
tb1 <- tb %>% select(
  salary.x.x, 
  salary.y, 
  salary.x, 
  year2.x, 
  utbregno, 
  perc_edu, 
  sum_edu, 
  salary.y.y, 
  perc_women, 
  sum_ing, 
  perc_eng)
  
tb2 <- tb1

tb3 <- tb2 %>% scale()  
  
d <- dist(tb3, method = "euclidean")

res.hc <- hclust(d, method = "ward.D2" )

grp <- cutree(res.hc, k = 2)

tb4 <- tb2 %>% mutate(grp)

model1 <- lm(perc_women ~ ., data = filter(tb4, grp == 1))

summary(model1) %>%	
  tidy() %>%
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Summary of model data for group 1, number of population with 3 years or more post-secondary education > 193000')

model2 <- lm(perc_women ~ ., data = filter(tb4, grp == 2))

summary(model2) %>%	
  tidy() %>%
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Women engineers and population with 3 years or more post-secondary education')

Anova(model2, type=2) %>% 
  tidy() %>% 
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Anova report from linear model fit') 

tb %>% mutate(grp) %>%
  ggplot(aes(x = salary.x.x, y = perc_edu, colour = region, size = year2.x)) +
  geom_point() +
  theme(legend.position="bottom") +
  facet_grid(. ~ grp) +
  labs(
    x = "Total R&D person years performed by employees of a higher education institution by region",
    y = "Per cent of engineers who are women"
  )
```