---
title: "Presentation 200121"
author: "Mikael Lundqvist"
date: "21 januari 2020"
output: revealjs::revealjs_presentation
self_contained: true
reveal_options:
autoSlide: 5000
loop: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



```{r , echo = FALSE}
library (tidyverse) 
library (broom) 
library (car)
library (polynom)
library (splines)
#install_github("ZheyuanLi/SplinesUtils")
library(SplinesUtils)
library(sjPlot)

readfile <- function (file1){  
  read_csv (file1, col_types = cols(), locale = readr::locale (encoding = "latin1"), na = c("..", "NA")) %>%  
    gather (starts_with("19"), starts_with("20"), key = "year", value = salary) %>%  
	drop_na() %>%  
    mutate (year_n = parse_number (year))
}

```

## Engineering professionals

```{r , echo = FALSE}
tb <- readfile("00000031.csv") %>% 
  rowwise() %>% 
  mutate(age_l = unlist(lapply(strsplit(substr(age, 1, 5), "-"), strtoi))[1]) %>%
  rowwise() %>% 
  mutate(age_h = unlist(lapply(strsplit(substr(age, 1, 5), "-"), strtoi))[2]) %>%
  mutate(age_n = (age_l + age_h) / 2) 
   
tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = salary, colour = age, shape=sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )
```

## Engineering professionals

```{r , echo = FALSE}
model <- lm (log(salary) ~ year_n + sex + bs(age_n, knots = c(30, 40, 50, 60)), data = tb)

tb <- bind_cols(tb, as_tibble(exp(predict(model, tb, interval = "confidence"))))

tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = fit, colour = age, shape=sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )
```

## Engineering professionals

```{r, echo = FALSE}
myspline <- RegBsplineAsPiecePoly(model, "bs(age_n, knots = c(30, 40, 50, 60))")

tibble(age_n = 21:65) %>%
  ggplot () + 
    geom_point (mapping = aes(x = age_n,y = predict(myspline, age_n))) +
  labs(
    x = "Age",
    y = "Salary"
  )
```

## Engineering professionals

```{r, echo = FALSE}
summod <- tidy(summary (model))

myfun <- function(x){
  return (predict(myspline, x, deriv = 1) + summod$estimate[2])
}

tibble(age_n = 21:65) %>%
  ggplot () + 
    geom_point (mapping = aes(x = age_n,y = myfun(age_n))) +
  labs(
    x = "Age",
    y = "Salary raise (%)"
  )
```

## Different occupational groups

```{r, echo = FALSE} 
tb <- readfile("000000D2.csv") %>%
  rowwise() %>%
  mutate(age_l = unlist(lapply(strsplit(substr(age, 1, 5), "-"), strtoi))[1]) %>%
  rowwise() %>%
  mutate(age_h = unlist(lapply(strsplit(substr(age, 1, 5), "-"), strtoi))[2]) %>%
  mutate(age_n = (age_l + age_h) / 2)

summary_table = 0
anova_table = 0

for (i in unique(tb$`occuptional  (SSYK 2012)`)){
  temp <- filter(tb, `occuptional  (SSYK 2012)` == i)
  if (dim(temp)[1] > 30){
    model <-lm (log(salary) ~ year_n + poly(age_n, 3), data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i))
  }
}

merge(summary_table, anova_table, by = "ssyk", all = TRUE) %>%
  filter (term.y == "poly(age_n, 3)") %>%
  filter (term.x == "year_n") %>%
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y)) +
    labs(
      x = "Increase in salaries (% / year)",
      y = "F-value for age"
    ) 
```

## Primary- and pre-school teachers

```{r, echo = FALSE} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "234 Primary- and pre-school teachers")
 
temp %>%
  ggplot () +
    geom_point (mapping = aes(x = year_n,y = salary, colour = age)) +
    facet_grid(. ~ sex) +   
    labs(
      x = "Year",
      y = "Salary (SEK/month)"
    ) 
```

## Primary- and pre-school teachers

```{r, echo = FALSE}
model <-lm (log(salary) ~ year_n + poly(age_n, 3, raw = T), data = temp)

summod <- tidy(summary (model))

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = age_n,y = age_n * summod$estimate[3] + summod$estimate[4] * age_n^2 + summod$estimate[5] * age_n^3)) +
    labs(
      x = "Age",
      y = "Salary"
    )
```

## Primary- and pre-school teachers

```{r, echo = FALSE}
pdx <- deriv(as.polynomial(c(0, summod$estimate[3], summod$estimate[4], summod$estimate[5])))

temp %>%
  ggplot () + 
    geom_point (mapping = aes(x = age_n, y = summod$estimate[2] + pdx[1] + pdx[2] * age_n + pdx[3] * age_n^2)) +
    labs(
      x = "Age",
      y = "Salary raise (%)"
    )
```

## Assemblers

```{r, echo = FALSE} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "821 Assemblers")

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = year_n,y = salary, colour = age)) +
    facet_grid(. ~ sex) +   
    labs(
      x = "Year",
      y = "Salary (SEK/month)"
    ) 
```

## Assemblers

```{r, echo = FALSE}
model <-lm (log(salary) ~ year_n + poly(age_n, 3, raw = T), data = temp)

summod <- tidy(summary (model))

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = age_n,y = age_n * summod$estimate[3] + summod$estimate[4] * age_n^2 + summod$estimate[5] * age_n^3)) +
    labs(
      x = "Age",
      y = "Salary"
    )
```

## Assemblers

```{r, echo = FALSE}
pdx <- deriv(as.polynomial(c(0, summod$estimate[3], summod$estimate[4], summod$estimate[5])))

temp %>%
  ggplot () + 
    geom_point (mapping = aes(x = age_n, y = summod$estimate[2] + pdx[1] + pdx[2] * age_n + pdx[3] * age_n^2)) +
    labs(
      x = "Age",
      y = "Salary raise (%)"
    )
```

## Financial and insurance managers

```{r, echo = FALSE} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "161 Financial and insurance managers")

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = year_n,y = salary, colour = age)) +
    facet_grid(. ~ sex) + 
      labs(
        x = "Year",
        y = "Salary (SEK/month)"
      ) 
```

## Financial and insurance managers

```{r, echo = FALSE}
model <- lm (log(salary) ~ year_n + poly(age_n, 3, raw = T), data = temp)

summod <- tidy(summary (model))

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = age_n,y = age_n * summod$estimate[3] + summod$estimate[4] * age_n^2 + summod$estimate[5] * age_n^3)) +
    labs(
      x = "Age",
      y = "Salary"
    )
```

## Financial and insurance managers

```{r, echo = FALSE}
pdx <- deriv(as.polynomial(c(0, summod$estimate[3], summod$estimate[4], summod$estimate[5])))

temp %>%
  ggplot () + 
    geom_point (mapping = aes(x = age_n, y = summod$estimate[2] + pdx[1] + pdx[2] * age_n + pdx[3] * age_n^2)) +
    labs(
      x = "Age",
      y = "Salary raise (%)"
    )
```

## Cashiers and related clerks

```{r, echo = FALSE} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "523 Cashiers and related clerks")
temp %>%
  ggplot () +
    geom_point (mapping = aes(x = year_n,y = salary, colour = age)) +
    facet_grid(. ~ sex) + 
    labs(
      x = "Year",
      y = "Salary (SEK/month)"
  )
```

## Cashiers and related clerks

```{r, echo = FALSE}
model <-lm (log(salary) ~ year_n + poly(age_n, 3, raw = T), data = temp)

summod <- tidy(summary (model))

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = age_n,y = age_n * summod$estimate[3] + summod$estimate[4] * age_n^2 + summod$estimate[5] * age_n^3)) +
    labs(
      x = "Age",
      y = "Salary"
    )
```

## Cashiers and related clerks

```{r, echo = FALSE}
pdx <- deriv(as.polynomial(c(0, summod$estimate[3], summod$estimate[4], summod$estimate[5])))

temp %>%
  ggplot () + 
    geom_point (mapping = aes(x = age_n, y = summod$estimate[2] + pdx[1] + pdx[2] * age_n + pdx[3] * age_n^2)) +
    labs(
      x = "Age",
      y = "Salary raise (%)"
    )
```

## Engineering professionals

```{r, echo = FALSE} 
tb <- readfile("000000CY.csv") %>% 
  filter(`occuptional  (SSYK 2012)` == "214 Engineering professionals") %>%       
  mutate(edulevel = `level of education`)

tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = salary, colour = `level of education`, shape=sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )

model <- lm (log(salary) ~ year_n + sex + edulevel, data = tb)

tb <- bind_cols(tb, as_tibble(exp(predict(model, tb, interval = "confidence"))))
```

## Engineering professionals

```{r, echo = FALSE}
tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = fit, colour = edulevel, shape = sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )
```

## Engineering professionals

```{r, echo = FALSE}
numedulevel <- read.csv("edulevel.csv") 
  
tbnum <- tb %>% 
    right_join(numedulevel, by = c("level of education" = "level.of.education")) %>%
  filter(!is.na(eduyears))

modelcont <- lm(log(salary) ~ bs(eduyears, knots = c(14)) + year_n + sex, data = tbnum)

tbnum <- bind_cols(tbnum, as_tibble(exp(predict(modelcont, tbnum, interval = "confidence"))))

tbnum %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = fit1, colour = eduyears, shape = sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )
```

## Engineering professionals

```{r, echo = FALSE}

contspline <- RegBsplineAsPiecePoly(modelcont, "bs(eduyears, knots = c(14))")

tibble(eduyears = 9:19) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = predict(contspline, eduyears))) +
  labs(
    x = "Years of education",
    y = "Salary"
  )
```

## Engineering professionals

```{r, echo = FALSE}

tibble(eduyears = 9:19) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = (exp(predict(contspline, eduyears, deriv = 1)) - 1) * 100)) +
  labs(
    x = "Years of education",
    y = "Salary difference (%)"
  )
```

## Different occupational groups

```{r, echo = FALSE} 
tb <- readfile("000000CY.csv")

numedulevel <- read.csv("edulevel.csv") 

tbnum <- tb %>% 
  right_join(numedulevel, by = c("level of education" = "level.of.education")) %>%
  filter(!is.na(eduyears))

summary_table = vector()
anova_table = vector()

for (i in unique(tbnum$`occuptional  (SSYK 2012)`)){
  temp <- filter(tbnum, `occuptional  (SSYK 2012)` == i)
  if (dim(temp)[1] > 30){
    model <- lm (log(salary) ~ year_n + eduyears, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i))
  }
}

merge(summary_table, anova_table, by = "ssyk", all = TRUE) %>%
  filter (term.y == "eduyears") %>%
  filter (term.x == "eduyears") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y)) +
    labs(
      x = "Increase in salaries (% / year of education)",
      y = "F-value for education"
    ) 
```

## Health care managers

```{r, echo = FALSE} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "151 Health care managers")

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = year_n,y = salary, colour = eduyears)) +
    facet_grid(. ~ sex) +   
    labs(
      x = "Year",
      y = "Salary (SEK/month)"
    ) 
```

## Health care managers

```{r, echo = FALSE}
modelcont <- lm (log(salary) ~ year_n + bs(eduyears, degree = 1), data = temp)

contspline <- RegBsplineAsPiecePoly(modelcont, "bs(eduyears, degree = 1)")

tibble(eduyears = seq(11, 19, by=0.1)) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = predict(contspline, eduyears))) +
  labs(
    x = "Years of education",
    y = "Salary"
  )
```

## Health care managers

```{r, echo = FALSE}
tibble(eduyears = seq(11, 19, by=0.1)) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = (exp(predict(contspline, eduyears, deriv = 1)) - 1) * 100)) +
  labs(
    x = "Years of education",
    y = "Salary difference (%)"
  )
```

## Creative and performing artists

```{r, echo = FALSE} 
temp <- tbnum %>% 
  filter(`occuptional  (SSYK 2012)` == "265 Creative and performing artists")

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = year_n,y = salary, colour = eduyears)) +
    facet_grid(. ~ sex) +   
    labs(
      x = "Year",
      y = "Salary (SEK/month)"
    ) 
```

## Creative and performing artists

```{r, echo = FALSE}
modelcont <- lm (log(salary) ~ year_n + bs(eduyears, degree = 1), data = temp)

contspline <- RegBsplineAsPiecePoly(modelcont, "bs(eduyears, degree = 1)")

tibble(eduyears = seq(11, 15, by=0.1)) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = predict(contspline, eduyears))) +
  labs(
    x = "Years of education",
    y = "Salary"
  )
```

## Creative and performing artists

```{r, echo = FALSE}
tibble(eduyears = seq(11, 15, by=0.1)) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = (exp(predict(contspline, eduyears, deriv = 1)) - 1) * 100)) +
  labs(
    x = "Years of education",
    y = "Salary difference (%)"
  )
```

## Engineering professionals

```{r, echo = FALSE}   
tb <- readfile("00000031.csv") %>% 
  rowwise() %>% 
  mutate(age_l = unlist(lapply(strsplit(substr(age, 1, 5), "-"), strtoi))[1]) %>%
  rowwise() %>% 
  mutate(age_h = unlist(lapply(strsplit(substr(age, 1, 5), "-"), strtoi))[2]) %>%
  mutate(age_n = (age_l + age_h) / 2) 
   
tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = salary, colour = age, shape = sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )
```

## Engineering professionals

```{r, echo = FALSE}  
model <- lm (log(salary) ~ year_n + sex * age, data = tb)

plot_model(model, type = "pred", terms = c("age", "sex"))
	
```

## Engineering professionals

```{r, echo = FALSE}
model <- lm (log(salary) ~ year_n + sex * bs(age_n, knots = c(35, 50)), data = tb)

plot_model(model, type = "pred", terms = c("age_n", "sex"))

```

## Engineering professionals

```{r, echo = FALSE}  

model <- lm (log(salary) ~ year_n * sex + bs(age_n, knots = c(35, 50)), data = tb)

plot_model(model, type = "pred", terms = c("year_n", "sex"))

```

## Engineering professionals

```{r, echo = FALSE} 
tb <- readfile("000000CY.csv") %>% 
  filter(`occuptional  (SSYK 2012)` == "214 Engineering professionals") %>%    
  mutate(edulevel = `level of education`)

numedulevel <- read.csv("edulevel.csv") 

tbnum <- tb %>% 
  right_join(numedulevel, by = c("level of education" = "level.of.education")) %>%
  filter(!is.na(eduyears))

tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = salary, colour = `level of education`, shape = sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )
```

## Engineering professionals

```{r, echo = FALSE} 
model <- lm(log(salary) ~ sex * edulevel + year_n, data = tbnum)

plot_model(model, type = "pred", terms = c("edulevel", "sex"))
```

## Engineering professionals

```{r, echo = FALSE} 

model <- lm(log(salary) ~ sex * bs(eduyears, knots = c(14)) + year_n, data = tbnum)

plot_model(model, type = "pred", terms = c("eduyears", "sex"))

```
