---
title: The significance of education on the salary of engineers in Sweden
author: Mikael Lundqvist
date: '2019-12-15'
slug: the-significance-of-education-on-the-salary-of-engineers-in-sweden
categories:
  - R
tags:
  - plot
  - R Markdown
  - regression
---

In my last posts I analysed the significance of experience for different occupational groups. In this post I will turn my interest to education. I will again start with engineers and see if I can expand my analysis to all occupational groups.


First, define libraries and functions.


```{r , echo = TRUE}
library (tidyverse) 
library (broom) 
library (car)
library (splines)
#install_github("ZheyuanLi/SplinesUtils")
library(SplinesUtils)

readfile <- function (file1){  
  read_csv (file1, col_types = cols(), locale = readr::locale (encoding = "latin1"), na = c("..", "NA")) %>%  
    gather (starts_with("19"), starts_with("20"), key = "year", value = salary) %>%  
	drop_na() %>%  
    mutate (year_n = parse_number (year))
}

``` 

The data table is downloaded from Statistics Sweden. It is saved as a comma-delimited file without heading, 000000CY.csv, http://www.statistikdatabasen.scb.se/pxweb/en/ssd/.

The table:
Average basic salary, monthly salary and women´s salary as a percentage of men´s salary by sector, occupational group (SSYK 2012), sex and educational level (SUN). Year 2014 - 2018
Monthly sakary
All sectors

When I analysed experience I converted the categorical predictor age, which was divided into age groups, into a continuous predictor.

```{r, echo = TRUE, fig.cap = ', Year 2014 - 2018'} 
tb <- readfile("000000CY.csv") %>% filter(`occuptional  (SSYK 2012)` == "214 Engineering professionals") %>% mutate(edulevel = `level of education`)

tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = salary, colour = `level of education`, shape=sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )

model <- lm (log(salary) ~ year_n + sex + edulevel, data = tb)

tb <- bind_cols(tb, as_tibble(exp(predict(model, tb, interval = "confidence"))))
```

The R-value from the Anova table for years is 40 (Pr(>F) < 2.2e-16), sufficient for rejecting the null hypothesis that education has no effect on the salary holding year as constant.
The adjusted R-squared value is 0,852 implying a good fit of the model.

```{r, echo = TRUE, fig.cap = 'Model fit, SSYK 214, Architects, engineers and related professionals, Year 2014 - 2018'}
tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = fit, colour = edulevel, shape = sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )
	
summary(model) %>%	
  tidy() %>%
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Summary from linear model fit')

summary(model)$r.squared   
  
Anova(model, type=2) %>% 
  tidy() %>% 
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Anova report from linear model fit')
```

How much does the different levels of education effect the salary. We can calculate the differences between the levels with Tukey's honest significant difference. All significant level differences are shown in the table below.

```{r, echo = TRUE, fig.cap = 'Model fit, SSYK 214, Architects, engineers and related professionals, Year 2014 - 2018'}
TukeyHSD(aov(log(salary) ~ sex + edu, data = tb1)) %>% 
  tidy() %>%
  filter(adj.p.value < 0.05) %>%
  arrange(desc(comparison)) %>% 
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Significant values from Tukey HSD sorted in falling level difference')
```