---
title: The significance of education on the salary in Sweden, a comparison between different occupational groups
author: "Mikael Lundqvist"
date: '2019-12-21'
slug: the-significance-of-education-on-the-salary-in-sweden-a-comparison-between-different-occupational-groups
tags:
- plot
- R Markdown
- regression
categories: R
---

In my last post, I found that education has a significant impact on the salary of engineers. Is the significance of education on wages unique to engineers or are there similar correlations in other occupational groups?

I will use the same model in principal as in my previous post to calculate the significance of education. I will not use sex as an explanatory variable since there are occupational groups that do not have enough data for both genders.  Searching through the different occupational groups I will fit education with a polynomial of degree one. I am interested in occupational groups where a longer education also results in higher salaries. Because of that, I will use the numerical approximation from my last post instead of using the categorical predictor. A polynomial of higher degree than one would result in a better fit but the problem with oscillation and overfitting made me settle for degree one. A straight line as a function also has the advantage that the average increase in salary for each education year is directly given from the model.

There are still occupational groups with too little data for regression analysis. More than 30 posts are necessary to fit both education and year.

The F-value from the Anova table is used as the single value to discriminate how much education and salary correlates. For exploratory analysis, the Anova value seems good enough. 

In the figure below I will also use the estimate for education to see how much the salaries are raised by education for the different occupational groups holding year as constant.


```{r , echo = TRUE}
library (tidyverse)
library (broom)
library (car)
library (splines)
#install_github("ZheyuanLi/SplinesUtils")
library (SplinesUtils)

readfile <- function (file1){read_csv (file1, col_types = cols(), locale = readr::locale (encoding = "latin1"), na = c("..", "NA")) %>%
  gather (starts_with("19"), starts_with("20"), key = "year", value = salary) %>%
  drop_na() %>%
  mutate (year_n = parse_number (year))
}
```


The data table is downloaded from Statistics Sweden. It is saved as a comma-delimited file without heading, 000000CY.csv, http://www.statistikdatabasen.scb.se/pxweb/en/ssd/.

The table:
Average basic salary, monthly salary and women´s salary as a percentage of men´s salary by sector, occupational group (SSYK 2012), sex and educational level (SUN). Year 2014 - 2018
Monthly salary
All sectors


```{r, echo = TRUE, fig.cap = 'The significance of education on the salary in Sweden, a comparison between different occupational groups, Year 2014 - 2018'} 
tb <- readfile("000000CY.csv")

numedulevel <- read.csv("edulevel.csv") 

numedulevel %>%
  knitr::kable(
  booktabs = TRUE,
  caption = 'Initial approach, length of education') 
  
tbnum <- tb %>% 
  right_join(numedulevel, by = c("level of education" = "level.of.education")) %>%
  filter(!is.na(eduyears))

summary_table = vector()
anova_table = vector()

for (i in unique(tbnum$`occuptional  (SSYK 2012)`)){
  temp <- filter(tbnum, `occuptional  (SSYK 2012)` == i)
  if (dim(temp)[1] > 30){
    model <- lm (log(salary) ~ year_n + eduyears, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i))
  }
}

merge(summary_table, anova_table, by = "ssyk", all = TRUE) %>%
  filter (term.y == "eduyears") %>%
  filter (term.x == "eduyears") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y)) +
    labs(
      x = "Increase in salaries (% / year of education)",
      y = "F-value for education"
    ) 
```


The table with all occupational groups sorted by Increase in salary in descending order.


```{r, echo = TRUE}
merge(summary_table, anova_table, by = "ssyk", all = TRUE) %>%
  filter (term.y == "eduyears") %>%
  filter (term.x == "eduyears") %>%
  select (ssyk, estimate, statistic.y) %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%
  rename (`F-value for education` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`Increase in salary`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (education) and the increase in salaries for each year of education')
```


Let's check what we have found.


```{r, echo = TRUE, fig.cap = 'Highest increase in salary, 151 Health care managers'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "151 Health care managers")

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = year_n,y = salary, colour = eduyears)) +
    facet_grid(. ~ sex) +   
    labs(
      x = "Year",
      y = "Salary (SEK/month)"
    ) 
```


```{r, echo = TRUE, fig.cap = 'Model fit, Health care managers, Correlation between education and salary'}
modelcont <- lm (log(salary) ~ year_n + bs(eduyears, degree = 1), data = temp)

contspline <- RegBsplineAsPiecePoly(modelcont, "bs(eduyears, degree = 1)")

tibble(eduyears = seq(11, 19, by=0.1)) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = predict(contspline, eduyears))) +
  labs(
    x = "Years of education",
    y = "Salary"
  )
```


```{r, echo = TRUE, fig.cap = 'Model fit, Health care managers, The derivative for educaton'}
tibble(eduyears = seq(11, 19, by=0.1)) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = (exp(predict(contspline, eduyears, deriv = 1)) - 1) * 100)) +
  labs(
    x = "Years of education",
    y = "Salary difference (%)"
  )
```


```{r, echo = TRUE, fig.cap = 'Lowest increase in salary, 265 Creative and performing artists'} 
temp <- tbnum %>% 
  filter(`occuptional  (SSYK 2012)` == "265 Creative and performing artists")

temp %>%
  ggplot () +
    geom_point (mapping = aes(x = year_n,y = salary, colour = eduyears)) +
    facet_grid(. ~ sex) +   
    labs(
      x = "Year",
      y = "Salary (SEK/month)"
    ) 
```


```{r, echo = TRUE, fig.cap = 'Model fit, Creative and performing artists, Correlation between education and salary'}
modelcont <- lm (log(salary) ~ year_n + bs(eduyears, degree = 1), data = temp)

contspline <- RegBsplineAsPiecePoly(modelcont, "bs(eduyears, degree = 1)")

tibble(eduyears = seq(11, 15, by=0.1)) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = predict(contspline, eduyears))) +
  labs(
    x = "Years of education",
    y = "Salary"
  )
```


```{r, echo = TRUE, fig.cap = 'Model fit, Creative and performing artists, The derivative for education'}
tibble(eduyears = seq(11, 15, by=0.1)) %>%
  ggplot () + 
    geom_point (mapping = aes(x = eduyears,y = (exp(predict(contspline, eduyears, deriv = 1)) - 1) * 100)) +
  labs(
    x = "Years of education",
    y = "Salary difference (%)"
  )
```