---
title: The significance of the region on the salary of engineers in Sweden
author: Mikael Lundqvist
date: '2020-01-29'
slug: the-significance-of-the-region-on-the-salary-of-engineers-in-sweden
categories:
  - R
tags:
  - plot
  - R Markdown
  - regression
---

So far I have analysed the effect of experience, education, gender and year on the salary of engineers in Sweden. In this post, I will have a look at the effect of the region on the salary of engineers in Sweden.

Statistics Sweden use NUTS (Nomenclature des Unités Territoriales Statistiques), which is the EU's hierarchical regional division, to specify the regions.

First, define libraries and functions.


```{r , echo = TRUE}
library (tidyverse) 
library (broom) 
library (car)
library (swemaps) # devtools::install_github('reinholdsson/swemaps')
library(sjPlot)
readfile <- function (file1){  
  read_csv (file1, col_types = cols(), locale = readr::locale (encoding = "latin1"), na = c("..", "NA")) %>%  
    gather (starts_with("19"), starts_with("20"), key = "year", value = salary) %>%  
	drop_na() %>%  
    mutate (year_n = parse_number (year))
}
nuts <- read.csv("nuts.csv") %>%
  mutate(NUTS2_sh = substr(NUTS2, 1, 4))

map_ln_n <- map_ln %>%
  mutate(lnkod_n = as.numeric(lnkod)) 
``` 


The data table is downloaded from Statistics Sweden. It is saved as a comma-delimited file without heading, 000000CG.csv, http://www.statistikdatabasen.scb.se/pxweb/en/ssd/.

The table: Average basic salary, monthly salary and women´s salary as a percentage of men´s salary by region, sector, occupational group (SSYK 2012) and sex. Year 2014 - 2018 Monthly salary All sectors

We expect that the region is an important factor in salaries. As a null hypothesis, we assume that the region is not related to the salary and examine if we can reject this hypothesis with the data from Statistics Sweden.


```{r, echo = TRUE, fig.cap = 'SSYK 214, Architects, engineers and related professionals, men, Year 2014 - 2018'} 
tb <- readfile ("000000CG.csv") %>%
  filter (`occuptional  (SSYK 2012)` == "214 Engineering professionals") %>% 
  left_join(nuts %>% distinct (NUTS2_en, NUTS2_sh), by = c("region" = "NUTS2_en")) 
  
tb_map <- readfile ("000000CG.csv") %>%
  filter (`occuptional  (SSYK 2012)` == "214 Engineering professionals") %>%
  left_join(nuts, by = c("region" = "NUTS2_en")) 

tb_map %>%
  filter (sex == "men") %>%
  right_join(map_ln_n, by = c("Länskod" = "lnkod_n")) %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'SSYK 214, Architects, engineers and related professionals, women, Year 2014 - 2018'} 
tb_map %>%
  filter (sex == "women") %>%
  right_join(map_ln_n, by = c("Länskod" = "lnkod_n")) %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'SSYK 214, Architects, engineers and related professionals, Year 2014 - 2018'} 
tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n, y = salary, colour = region, shape=sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )

```


The F-value from the Anova table for the region is 40 (Pr(>F) < 2.2e-16), sufficient for rejecting the null hypothesis that the region has no effect on the salary holding year as constant. The adjusted R-squared value is 0,882 implying a good fit of the model.


```{r, echo = TRUE, fig.cap = 'Model fit, SSYK 214, Architects, engineers and related professionals, Year 2014 - 2018'} 

model <-lm (log(salary) ~ year_n + sex + region , data = tb)

tb <- bind_cols(tb, as_tibble(exp(predict(model, tb, interval = "confidence"))))

tb %>%
  ggplot () +  
    geom_point (mapping = aes(x = year_n,y = fit, colour = region, shape=sex)) + 
  labs(
    x = "Year",
    y = "Salary (SEK/month)"
  )

summary(model) %>%  
  tidy() %>%
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Summary from linear model fit')

summary(model)$r.squared

Anova(model, type=2) %>% 
  tidy() %>% 
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Anova report from linear model fit')
```


Let's check what we have found. 


```{r, echo = TRUE, fig.cap = 'SSYK 214, Architects, engineers and related professionals, Year 2014 - 2018'} 
model <-lm (log(salary) ~ year_n + sex + NUTS2_sh , data = tb) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh"))

plot(model, which = 1)

tb[38,]

tb[27,]

tb[5,]
```


Also examine the interaction between gender and region. The F-value from the Anova table for the interaction is 2,7 (Pr(>F) < 0.017) implying a relation to 95 % significance.


```{r, echo = TRUE, fig.cap = 'SSYK 214, Architects, engineers and related professionals, Year 2014 - 2018'} 
model <-lm (log(salary) ~ year_n + sex * NUTS2_sh , data = tb) 

summary(model) %>%  
  tidy() %>%
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Summary from linear model fit')

summary(model)$r.squared

Anova(model, type=2) %>% 
  tidy() %>% 
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Anova report from linear model fit')
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "sex"))

plot(model, which = 1)

tb[37,]
```


And the interaction between year and region. The F-value from the Anova table for the region is 0,57 (Pr(>F) < 0,77) implying no significant relation.


```{r, echo = TRUE, fig.cap = 'SSYK 214, Architects, engineers and related professionals, Year 2014 - 2018'} 
model <-lm (log(salary) ~ year_n * NUTS2_sh + sex , data = tb) 
 
summary(model) %>%  
  tidy() %>%
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Summary from linear model fit')

summary(model)$r.squared

Anova(model, type=2) %>% 
  tidy() %>% 
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Anova report from linear model fit')

plot_model(model, type = "pred", terms = c("NUTS2_sh", "year_n"))

plot(model, which = 1)
```


Finally the interaction between gender, year and region, the only significant interaction is between the region and gender, F-value from the Anova table for the interaction is 2,4 (Pr(>F) < 0.033)  


```{r, echo = TRUE, fig.cap = 'SSYK 214, Architects, engineers and related professionals, Year 2014 - 2018'} 
model <-lm (log(salary) ~ year_n * NUTS2_sh * sex , data = tb) 
 
summary(model) %>%  
  tidy() %>%
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Summary from linear model fit')

summary(model)$r.squared

Anova(model, type=2) %>% 
  tidy() %>% 
  knitr::kable( 
  booktabs = TRUE,
  caption = 'Anova report from linear model fit')

plot_model(model, type = "pred", terms = c("NUTS2_sh", "year_n", "sex"))

plot(model, which = 1)
```