---
title: The significance of the sector on the salary in Sweden, a comparison between different occupational groups
author: Mikael Lundqvist
date: '2020-02-19'
slug: the-significance-of-the-sector-on-the-salary-in-sweden-a-comparison-between-different-occupational-groups
categories:
  - R
tags:
  - plot
  - R Markdown
  - regression
---

In my last post, I found that the sector has a significant impact on the salary of engineers. Is the significance of the sector unique to engineers or are there similar correlations in other occupational groups?

Statistics Sweden use NUTS (Nomenclature des Unités Territoriales Statistiques), which is the EU’s hierarchical regional division, to specify the regions.

The F-value from the Anova table is used as the single value to discriminate how much the region and salary correlates. For exploratory analysis, the Anova value seems good enough.

First, define libraries and functions.


```{r , echo = TRUE}
library (tidyverse) 
library (broom) 
library (car)
library (swemaps) # devtools::install_github('reinholdsson/swemaps')
library(sjPlot)
readfile <- function (file1){  
  read_csv (file1, col_types = cols(), locale = readr::locale (encoding = "latin1"), na = c("..", "NA")) %>%  
    gather (starts_with("19"), starts_with("20"), key = "year", value = salary) %>%  
	drop_na() %>%  
    mutate (year_n = parse_number (year))
}
nuts <-read_csv ("nuts.csv", col_types = cols(), locale = readr::locale (encoding = "latin1"), na = c("..", "NA")) %>%
  mutate(NUTS2_sh = substr(NUTS2, 1, 4))
nuts %>% 
  distinct (NUTS2_en) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Nomenclature des Unités Territoriales Statistiques (NUTS)')

map_ln_n <- map_ln %>%
  mutate(lnkod_n = as.numeric(lnkod)) 
``` 


The data table is downloaded from Statistics Sweden. It is saved as a comma-delimited file without heading, 000000CG.csv, http://www.statistikdatabasen.scb.se/pxweb/en/ssd/.

I have renamed the file to 000000CG_sector.csv because the filename 000000CG.csv was used in a previous post. 

The table: Average basic salary, monthly salary and women´s salary as a percentage of men´s salary by region, sector, occupational group (SSYK 2012) and sex. Year 2014 - 2018 Monthly salary 1-3 public sector 4-5 private sector

Only 17 occupational groups have employees in both the public and the private sector in all regions and both genders.

In the plot and tables, you can also find information on how the increase in salaries per year for each occupational group is affected when the interactions are taken into account.


```{r, echo = TRUE, fig.cap = 'The significance of the sector on the salary in Sweden, a comparison between different occupational groups, Year 2014 - 2018'} 
tb <- readfile ("000000CG_sector.csv") %>%
  left_join(nuts %>% distinct (NUTS2_en, NUTS2_sh), by = c("region" = "NUTS2_en")) 
  
tb_map <- readfile ("000000CG_sector.csv") %>%
  left_join(nuts, by = c("region" = "NUTS2_en")) %>%  
  right_join(map_ln_n, by = c("Länskod" = "lnkod_n"))


summary_table = vector()
anova_table = vector()
for (i in unique(tb$`occuptional  (SSYK 2012)`)){
  temp <- filter(tb, `occuptional  (SSYK 2012)` == i)
  if (dim(temp)[1] > 150){
    model <- lm(log(salary) ~ region + sex + year_n + sector, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "none"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "none"))  
  
    model <- lm(log(salary) ~ region + sex + year_n * sector, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "sector and year"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "sector and year"))
    
    model <- lm(log(salary) ~ region + year_n + sex * sector, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "sex and sector"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "sex and sector"))   
    
    model <- lm(log(salary) ~ region * sector + year_n + sex, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "region and sector"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "region and sector"))
    
    model <- lm(log(salary) ~ region * sector * year_n * sex, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "region, sector, year and sex"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "region, sector, year and sex"))    
  }
}

anova_table <- anova_table %>% rowwise() %>% mutate(contcol = str_count(term, ":")) 

summary_table <- summary_table %>% rowwise() %>% mutate(contcol = str_count(term, ":"))

merge(summary_table, anova_table, by = c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (term.y == "sector") %>%    
  filter (interaction == "none") %>%
  
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y, colour = interaction)) +
    labs(
      x = "Increase in salaries (% / year)",
      y = "F-value for sector"
    )  	
```


```{r, echo = TRUE, fig.cap = 'The significance of the interaction between sector, region, year and sex on the salary in Sweden, a comparison between different occupational groups, Year 2014 - 2018'}
merge(summary_table, anova_table, by = c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%    
  # only look at the interactions between all four variables in the case with interaction region, sector, year and sex
  filter (!(contcol.y < 3 & interaction == "region, sector, year and sex")) %>% 
  
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y, colour = interaction)) +
    labs(
      x = "Increase in salaries (% / year)",
      y = "F-value for interaction"
    ) 	
```


The tables with all occupational groups sorted by F-value in descending order.


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (term.y == "sector") %>%   
  filter (interaction == "none") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sector) and the yearly increase in salaries')
```


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "sector and year") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sector and year) and the yearly increase in salaries')
```


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "sex and sector") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sex and sector) and the yearly increase in salaries')
```


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "region and sector") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (region and sector) and the yearly increase in salaries')
```


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 1) %>%   
  filter (interaction == "region and sector") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (region, year and sex) and the yearly increase in salaries')
```


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 1) %>%   
  filter (interaction == "region, sector, year and sex") %>%
  filter (!(contcol.y < 3 & interaction == "region, sector, year and sex")) %>%  
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (region, year and sex) and the yearly increase in salaries')
```


Let's check what we have found.


```{r, echo = TRUE, fig.cap = 'Highest F-value sector, Newspaper distributors, janitors and other service workers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "962 Newspaper distributors, janitors and other service workers")
 
model <-lm (log(salary) ~ year_n + sex + NUTS2_sh + sector, data = temp) 
 
plot_model(model, type = "pred", terms = c("sector"))

tb_map %>%
    filter(`occuptional  (SSYK 2012)` == "962 Newspaper distributors, janitors and other service workers") %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ sector) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value sector, Insurance advisers, sales and purchasing agents'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "332 Insurance advisers, sales and purchasing agents")
 
model <-lm (log(salary) ~ year_n + sex + NUTS2_sh + sector, data = temp) 
 
plot_model(model, type = "pred", terms = c("sector"))
```


```{r, echo = TRUE, fig.cap = 'Highest F-value interaction gender and sector, Accountants, financial analysts and fund managers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "241 Accountants, financial analysts and fund managers")
 
model <-lm (log(salary) ~ year_n + sex * sector + NUTS2_sh, data = temp) 
 
plot_model(model, type = "pred", terms = c("sector", "sex"))

tb_map %>%
    filter(`occuptional  (SSYK 2012)` == "241 Accountants, financial analysts and fund managers") %>%
  filter (sector == "1-3 public sector") %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ sex) + 
    coord_equal() 

tb_map %>%
    filter(`occuptional  (SSYK 2012)` == "241 Accountants, financial analysts and fund managers") %>%
  filter (sector == "4-5 private sector") %>%  
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ sex) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction gender and sector, Client information clerks'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "422 Client information clerks")
 
model <-lm (log(salary) ~ year_n + sex * sector + NUTS2_sh, data = temp) 
 
plot_model(model, type = "pred", terms = c("sector", "sex"))
```


```{r, echo = TRUE, fig.cap = 'Highest F-value interaction year and sector, Client information clerks'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "422 Client information clerks")
 
model <-lm (log(salary) ~ year_n * sector + NUTS2_sh + sex , data = temp) 
 
plot_model(model, type = "pred", terms = c("year_n", "sector"))

tb_map %>%
  filter(`occuptional  (SSYK 2012)` == "422 Client information clerks") %>%
  filter (sector == "1-3 public sector") %>%  
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 

tb_map %>%
  filter(`occuptional  (SSYK 2012)` == "422 Client information clerks") %>%
  filter (sector == "4-5 private sector") %>%  
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction year and sector, Financial and accounting associate professionals'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "331 Financial and accounting associate professionals")
 
model <-lm (log(salary) ~ year_n * sector + NUTS2_sh + sex , data = temp) 
 
plot_model(model, type = "pred", terms = c("year_n", "sector"))
```


```{r, echo = TRUE, fig.cap = 'Highest F-value interaction region and sector, Stores and transport clerks'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "432 Stores and transport clerks")
 
model <-lm (log(salary) ~ year_n + sector * NUTS2_sh + sex , data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "sector"))

tb_map %>%
  filter(`occuptional  (SSYK 2012)` == "432 Stores and transport clerks") %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ sector) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction sector and region, Fast-food workers, food preparation assistants'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "941 Fast-food workers, food preparation assistants")
 
model <-lm (log(salary) ~ year_n + sector * NUTS2_sh + sex , data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "sector"))
```


```{r, echo = TRUE, fig.cap = 'Highest F-value interaction sector, region, year and gender, ICT operations and user support technicians'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "351 ICT operations and user support technicians")
 
model <-lm (log(salary) ~ year_n * NUTS2_sh * sex * sector, data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "year_n", "sector",  "sex"))

tb_map %>%
  filter(`occuptional  (SSYK 2012)` == "351 ICT operations and user support technicians") %>%
  filter (sector == "1-3 public sector") %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 

tb_map %>%
    filter(`occuptional  (SSYK 2012)` == "351 ICT operations and user support technicians") %>%
  filter (sector == "4-5 private sector") %>%  
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction sector, region, year and gender, Client information clerks'} 
  temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "422 Client information clerks")
 
model <-lm (log(salary) ~ year_n * NUTS2_sh * sex * sector, data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "year_n", "sector",  "sex"))
```
