---
title: The significance of the region on the salary in Sweden, a comparison between different occupational groups
author: Mikael Lundqvist
date: '2020-02-01'
slug: the-significance-of-the-region-on-the-salary-in-sweden-a-comparison-between-different-occupational-groups
categories:
  - R
tags:
  - plot
  - R Markdown
  - regression
---

In my last post, I found that the region has a significant impact on the salary of engineers. Is the significance of the region unique to engineers or are there similar correlations in other occupational groups?

Statistics Sweden use NUTS (Nomenclature des Unités Territoriales Statistiques), which is the EU’s hierarchical regional division, to specify the regions.

The F-value from the Anova table is used as the single value to discriminate how much the region and salary correlates. For exploratory analysis, the Anova value seems good enough.

First, define libraries and functions.


```{r , echo = TRUE}
library (tidyverse) 
library (broom) 
library (car)
library (swemaps) # devtools::install_github('reinholdsson/swemaps')
library(sjPlot)
readfile <- function (file1){  
  read_csv (file1, col_types = cols(), locale = readr::locale (encoding = "latin1"), na = c("..", "NA")) %>%  
    gather (starts_with("19"), starts_with("20"), key = "year", value = salary) %>%  
	drop_na() %>%  
    mutate (year_n = parse_number (year))
}
nuts <- read.csv("nuts.csv") %>%
  mutate(NUTS2_sh = substr(NUTS2, 1, 4))
nuts %>% 
  distinct (NUTS2_en) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Nomenclature des Unités Territoriales Statistiques (NUTS)')

map_ln_n <- map_ln %>%
  mutate(lnkod_n = as.numeric(lnkod)) 
``` 


The data table is downloaded from Statistics Sweden. It is saved as a comma-delimited file without heading, 000000CG.csv, http://www.statistikdatabasen.scb.se/pxweb/en/ssd/.

The table: Average basic salary, monthly salary and women´s salary as a percentage of men´s salary by region, sector, occupational group (SSYK 2012) and sex. Year 2014 - 2018 Monthly salary All sectors

In the plot and tables, you can also find information on how the increase in salaries per year for each occupational group is affected when the interactions are taken into account.


```{r, echo = TRUE, fig.cap = 'The significance of the region on the salary in Sweden, a comparison between different occupational groups, Year 2014 - 2018'} 
tb <- readfile ("000000CG.csv") %>%
  left_join(nuts %>% distinct (NUTS2_en, NUTS2_sh), by = c("region" = "NUTS2_en")) 
  
tb_map <- readfile ("000000CG.csv") %>%
  left_join(nuts, by = c("region" = "NUTS2_en")) %>%  
  right_join(map_ln_n, by = c("Länskod" = "lnkod_n"))


summary_table = vector()
anova_table = vector()
for (i in unique(tb$`occuptional  (SSYK 2012)`)){
  temp <- filter(tb, `occuptional  (SSYK 2012)` == i)
  if (dim(temp)[1] > 75){
    model <- lm(log(salary) ~ region + sex + year_n, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "none"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "none"))  
  
    model <- lm(log(salary) ~ region * sex + year_n, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "region and sex"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "region and sex"))
    
    model <- lm(log(salary) ~ region * year_n + sex, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "region and year"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "region and year"))   
    
    model <- lm(log(salary) ~ region * year_n * sex, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "region, year and sex"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "region, year and sex"))     
  }
}

anova_table <- anova_table %>% rowwise() %>% mutate(contcol = str_count(term, ":")) 

summary_table <- summary_table %>% rowwise() %>% mutate(contcol = str_count(term, ":"))

merge(summary_table, anova_table, by = c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (term.y == "region") %>%    
  filter (interaction == "none") %>%
  
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y, colour = interaction)) +
    labs(
      x = "Increase in salaries (% / year)",
      y = "F-value for interaction"
    ) 	
```


```{r, echo = TRUE, fig.cap = 'The significance of the interaction between region, year and sex on the salary in Sweden, a comparison between different occupational groups, Year 2014 - 2018'}
merge(summary_table, anova_table, by = c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%    
  # only look at the interactions between all three variables in the case with interaction region, year and sex
  filter (!(contcol.y == 1 & interaction == "region, year and sex")) %>% 
  
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y, colour = interaction)) +
    labs(
      x = "Increase in salaries (% / year)",
      y = "F-value for interaction"
    ) 	

```


The tables with all occupational groups sorted by F-value in descending order.


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (term.y == "region") %>%   
  filter (interaction == "none") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (region) and the yearly increase in salaries')
```


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "region and sex") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (region and sex) and the yearly increase in salaries')
```


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "region and year") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (region and year) and the yearly increase in salaries')
```


```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 1) %>%   
  filter (interaction == "region, year and sex") %>%
  filter (!(contcol.y == 1 & interaction == "region, year and sex")) %>%   
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (region, year and sex) and the yearly increase in salaries')
```


Let's check what we have found.


```{r, echo = TRUE, fig.cap = 'Highest F-value region, Primary- and pre-school teachers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "234 Primary- and pre-school teachers")
 
model <-lm (log(salary) ~ year_n + sex + NUTS2_sh, data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh"))

tb_map %>%
    filter(`occuptional  (SSYK 2012)` == "234 Primary- and pre-school teachers") %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value region, Cleaners and helpers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "911 Cleaners and helpers")
 
model <-lm (log(salary) ~ year_n + sex + NUTS2_sh, data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh"))
```


```{r, echo = TRUE, fig.cap = 'Highest F-value interaction gender and region, Financial and accounting associate professionals'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "331 Financial and accounting associate professionals")
 
model <-lm (log(salary) ~ year_n + sex * NUTS2_sh, data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "sex"))

tb_map %>%
    filter(`occuptional  (SSYK 2012)` == "331 Financial and accounting associate professionals") %>%
  filter (sex == "men") %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 

tb_map %>%
    filter(`occuptional  (SSYK 2012)` == "331 Financial and accounting associate professionals") %>%
  filter (sex == "women") %>%  
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction gender and region, Nursing professionals'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "222 Nursing professionals")
 
model <-lm (log(salary) ~ year_n + sex * NUTS2_sh, data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "sex"))
```


```{r, echo = TRUE, fig.cap = 'Highest F-value interaction year and region, Nursing professionals'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "222 Nursing professionals")
 
model <-lm (log(salary) ~ year_n * NUTS2_sh + sex , data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "year_n"))

tb_map %>%
  filter(`occuptional  (SSYK 2012)` == "222 Nursing professionals") %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 

```


```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction year and region, Other surveillance and security workers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "541 Other surveillance and security workers")
 
model <-lm (log(salary) ~ year_n * NUTS2_sh + sex , data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "year_n"))
```


```{r, echo = TRUE, fig.cap = 'Highest F-value interaction region, year and gender, Child care workers and teachers aides'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "531 Child care workers and teachers aides")
 
model <-lm (log(salary) ~ year_n * NUTS2_sh * sex , data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "year_n", "sex"))

tb_map %>%
  filter(`occuptional  (SSYK 2012)` == "531 Child care workers and teachers aides") %>%
  filter (sex == "men") %>%
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 

tb_map %>%
    filter(`occuptional  (SSYK 2012)` == "531 Child care workers and teachers aides") %>%
  filter (sex == "women") %>%  
  ggplot() +
    geom_polygon(mapping = aes(x = ggplot_long, y = ggplot_lat, group = lnkod, fill = salary)) +
      facet_grid(. ~ year) + 
    coord_equal() 
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction region, year and gender, Police officers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "336 Police officers")
 
model <-lm (log(salary) ~ year_n * NUTS2_sh * sex , data = temp) 
 
plot_model(model, type = "pred", terms = c("NUTS2_sh", "year_n", "sex"))
```