---
title: The significance of gender on the salary in Sweden, a comparison between different occupational groups
author: Mikael Lundqvist
date: '2020-01-27'
slug: the-significance-of-gender-on-the-salary-in-sweden-a-comparison-between-different-occupational-groups
categories:
  - R
tags:
  - plot
  - R Markdown
  - regression
---

In my last post, I found that the interaction between the different predictors has a significant impact on the salary of engineers. Is the significance of the interactions on wages unique to engineers or are there similar correlations in other occupational groups?

We start by examining the interaction between age, year, gender. 

The F-value from the Anova table is used as the single value to discriminate how much education and salary correlates. For exploratory analysis, the Anova value seems good enough.

First, define libraries and functions.


```{r , echo = TRUE}
library (tidyverse) 
library (broom) 
library (car)
library(sjPlot)
readfile <- function (file1){  
  read_csv (file1, col_types = cols(), locale = readr::locale (encoding = "latin1"), na = c("..", "NA")) %>%  
    gather (starts_with("19"), starts_with("20"), key = "year", value = salary) %>%  
	drop_na() %>%  
    mutate (year_n = parse_number (year))
}
``` 

The data table is downloaded from Statistics Sweden. It is saved as a comma-delimited file without heading, 000000D2.csv, http://www.statistikdatabasen.scb.se/pxweb/en/ssd/.

The table: Average basic salary, monthly salary and women´s salary as a percentage of men´s salary by sector, occupational group (SSYK 2012), sex and age. Year 2014 - 2018 Monthly salary All sectors

I will use a continuous predictor, a polynomial of degree three, to fit the contribution of age to the salary.

In the plot and tables, you can also find information on how the increase in salaries per year for each occupational group is affected when the interactions are taken into account.
 
 
```{r, echo = TRUE, fig.cap = 'The significance of the interaction between age, year, sex on the salary in Sweden, a comparison between different occupational groups, Year 2014 - 2018'} 
tb <- readfile("000000D2.csv") %>%
  rowwise() %>%
  mutate(age_l = unlist(lapply(strsplit(substr(age, 1, 5), "-"), strtoi))[1]) %>%
  rowwise() %>%
  mutate(age_h = unlist(lapply(strsplit(substr(age, 1, 5), "-"), strtoi))[2]) %>%
  mutate(age_n = (age_l + age_h) / 2)

summary_table = vector()
anova_table = vector()
for (i in unique(tb$`occuptional  (SSYK 2012)`)){
  temp <- filter(tb, `occuptional  (SSYK 2012)` == i)
  if (dim(temp)[1] > 30){
    model <-lm (log(salary) ~ year_n + sex * poly(age_n, 3), data = temp)
    summary_table <- bind_rows (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "sex and age"))
    anova_table <- bind_rows (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "sex and age"))

    model <-lm (log(salary) ~ year_n * sex + poly(age_n, 3), data = temp)
    summary_table <- bind_rows (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "sex and year"))
    anova_table <- bind_rows (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "sex and year"))

    model <-lm (log(salary) ~ sex + year_n * poly(age_n, 3), data = temp)
    summary_table <- bind_rows (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "year and age"))
    anova_table <- bind_rows (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "year and age"))

    model <-lm (log(salary) ~ sex * year_n * poly(age_n, 3), data = temp)
    summary_table <- bind_rows (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "sex, year and age"))
    anova_table <- bind_rows (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "sex, year and age"))
	}
}

anova_table <- anova_table %>% rowwise() %>% mutate(contcol = str_count(term, ":")) 

summary_table <- summary_table %>% rowwise() %>% mutate(contcol = str_count(term, ":"))

merge(summary_table, anova_table, by = c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%    
  # only look at the interactions between all three variables
  filter (!(contcol.y == 1 & interaction == "sex, year and age")) %>% 
  
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y, colour = interaction)) +
    labs(
      x = "Increase in salaries (% / year)",
      y = "F-value for interaction"
    ) 	
```

The tables with all occupational groups sorted by F-value in descending order.

```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "sex and age") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sex and age) and the yearly increase in salaries')
```

```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "sex and year") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sex and year) and the yearly increase in salaries')
```

```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "year and age") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (year and age) and the yearly increase in salaries')
```

```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 1) %>%   
  filter (interaction == "sex, year and age") %>%
  filter (!(contcol.y == 1 & interaction == "sex, year and age")) %>%   
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sex, year and age) and the yearly increase in salaries')
```

Let's check what we have found.

```{r, echo = TRUE, fig.cap = 'Highest F-value interaction sex and age, Tax and related government associate professionals'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "335 Tax and related government associate professionals")
 
model <-lm (log(salary) ~ year_n + sex * poly(age_n, 3), data = temp) 
 
plot_model(model, type = "pred", terms = c("age_n", "sex"))
```


```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction sex and age, Social work and religious associate professionals'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "341 Social work and religious associate professionals")
 
model <-lm (log(salary) ~ year_n + sex * poly(age_n, 3), data = temp) 
 
plot_model(model, type = "pred", terms = c("age_n", "sex"))
```

```{r, echo = TRUE, fig.cap = 'Highest F-value interaction sex and year, Information and communications technology service managers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "131 Information and communications technology service managers")
 
model <-lm (log(salary) ~ year_n * sex + poly(age_n, 3), data = temp) 
 
plot_model(model, type = "pred", terms = c("year_n", "sex"))
```

```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction sex and year, Sales and marketing managers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "125 Sales and marketing managers")
 
model <-lm (log(salary) ~ year_n * sex + poly(age_n, 3), data = temp) 
 
plot_model(model, type = "pred", terms = c("year_n", "sex"))
```

```{r, echo = TRUE, fig.cap = 'Highest F-value interaction year and age, Building caretakers and related workers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "515 Building caretakers and related workers")
 
model <-lm (log(salary) ~ sex + year_n * poly(age_n, 3), data = temp) 
 
plot_model(model, type = "pred", terms = c("age_n", "year_n"))
```

```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction year and age, Administration and service managers not elsewhere classified'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "129 Administration and service managers not elsewhere classified")
 
model <-lm (log(salary) ~ sex + year_n * poly(age_n, 3), data = temp) 
 
plot_model(model, type = "pred", terms = c("age_n", "year_n"))
```

```{r, echo = TRUE, fig.cap = 'Highest F-value interaction sex, year and age, Financial and insurance managers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "161 Financial and insurance managers")
 
model <-lm (log(salary) ~ sex * year_n * poly(age_n, 3), data = temp) 
 
plot_model(model, type = "pred", terms = c("age_n", "year_n", "sex"))
```

```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction sex, year and age, ICT architects, systems analysts and test managers'} 
temp <- tb %>%
  filter(`occuptional  (SSYK 2012)` == "251 ICT architects, systems analysts and test managers")
 
model <-lm (log(salary) ~ sex * year_n * poly(age_n, 3), data = temp) 
 
plot_model(model, type = "pred", terms = c("age_n", "year_n", "sex"))
```

We proceed by examining the interaction between education, year, gender. 

The data table is downloaded from Statistics Sweden. It is saved as a comma-delimited file without heading, 000000CY.csv, http://www.statistikdatabasen.scb.se/pxweb/en/ssd/.

The table:
Average basic salary, monthly salary and women´s salary as a percentage of men´s salary by sector, occupational group (SSYK 2012), sex and educational level (SUN). Year 2014 - 2018
Monthly salary
All sectors

I will use a categorical predictor to fit the contribution of education to the salary.

In the plot and tables, you can also find information on how the increase in salaries per year for each occupational group is affected when the interactions are taken into account.


```{r, echo = TRUE, fig.cap = 'The significance of education, year, sex and the interaction between them on the salary in Sweden, a comparison between different occupational groups, Year 2014 - 2018'} 
tb <- readfile("000000CY.csv") %>% 
  mutate(edulevel = `level of education`)

numedulevel <- read.csv("edulevel.csv") 
  
tbnum <- tb %>% 
  right_join(numedulevel, by = c("level of education" = "level.of.education")) %>%
  filter(!is.na(eduyears)) %>%
  mutate(eduyears = factor(eduyears))
  
summary_table = vector()
anova_table = vector()
for (i in unique(tbnum$`occuptional  (SSYK 2012)`)){
  temp <- filter(tbnum, `occuptional  (SSYK 2012)` == i)
  if (dim(temp)[1] > 30){
    model <-lm (log(salary) ~ year_n + sex * edulevel, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "sex and edulevel"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "sex and edulevel"))
	
	model <-lm (log(salary) ~ year_n * sex + edulevel, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "sex and year"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "sex and year"))

    model <-lm (log(salary) ~ sex + year_n * edulevel, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "year and edulevel"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "year and edulevel"))
	
    model <-lm (log(salary) ~ year_n * sex * edulevel, data = temp)
    summary_table <- rbind (summary_table, mutate (tidy (summary (model)), ssyk = i, interaction = "sex, year and edulevel"))
    anova_table <- rbind (anova_table, mutate (tidy (Anova (model, type = 2)), ssyk = i, interaction = "sex, year and edulevel"))	
	}
}  

anova_table <- anova_table %>% rowwise() %>% mutate(contcol = str_count(term, ":"))  

summary_table <- summary_table %>% rowwise() %>% mutate(contcol = str_count(term, ":"))

merge(summary_table, anova_table, by = c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%  
  filter (!(contcol.y == 1 & interaction == "sex, year and edulevel")) %>% 
  mutate (estimate = (exp(estimate) - 1) * 100) %>%
  ggplot () +
    geom_point (mapping = aes(x = estimate, y = statistic.y, colour = interaction)) +
    labs(
      x = "Increase in salaries (% / year)",
      y = "F-value for education"
    ) 
```  

The table with all occupational groups sorted by F-value in descending order.

```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "sex and edulevel") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sex and edulevel) and the yearly increase in salaries')
```

```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "sex and year") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sex and year) and the yearly increase in salaries')
```

```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 0) %>%   
  filter (interaction == "year and edulevel") %>%
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (year and edulevel) and the yearly increase in salaries')
```

```{r, echo = TRUE}
merge(summary_table, anova_table, c("ssyk", "interaction"), all = TRUE) %>%
  filter (term.x == "year_n") %>%
  filter (contcol.y > 1) %>%   
  filter (interaction == "sex, year and edulevel") %>%
  filter (!(contcol.y == 1 & interaction == "sex, year and edulevel")) %>% 
  mutate (estimate = (exp(estimate) - 1) * 100) %>%  
  select (ssyk, estimate, statistic.y, interaction) %>%
  rename (`F-value for age` = statistic.y) %>%
  rename (`Increase in salary` = estimate) %>%
  arrange (desc (`F-value for age`)) %>%
  knitr::kable(
    booktabs = TRUE,
    caption = 'Correlation for F-value (sex, year and edulevel) and the yearly increase in salaries')
```

```{r, echo = TRUE, fig.cap = 'Highest F-value interaction sex and education level, University and higher education teachers'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "231 University and higher education teachers") 
 
model <-lm (log(salary) ~ year_n + sex * eduyears, data = temp)
 
plot_model(model, type = "pred", terms = c("eduyears", "sex"))
```

```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction sex and education level, Machine operators, rubber, plastic and paper products'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "814 Machine operators, rubber, plastic and paper products") 
 
model <-lm (log(salary) ~ year_n + sex * eduyears, data = temp)
 
plot_model(model, type = "pred", terms = c("eduyears", "sex"))
```

```{r, echo = TRUE, fig.cap = 'Highest F-value interaction year and education level, Attendants, personal assistants and related workers'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "534 Attendants, personal assistants and related workers") 
 
model <-lm (log(salary) ~ sex + year_n * eduyears, data = temp)
 
plot_model(model, type = "pred", terms = c("year_n", "eduyears"))
```

```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction year and education level, Machine operators, rubber, plastic and paper products'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "814 Machine operators, rubber, plastic and paper products") 
 
model <-lm (log(salary) ~ sex + year_n * eduyears, data = temp)
 
plot_model(model, type = "pred", terms = c("year_n", "eduyears"))
```

```{r, echo = TRUE, fig.cap = 'Highest F-value interaction sex and year, Museum curators and librarians and related professionals'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "262 Museum curators and librarians and related professionals") 
 
model <-lm (log(salary) ~ sex * year_n + edulevel, data = temp)
 
plot_model(model, type = "pred", terms = c("year_n", "sex"))
```

```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction sex and year, Recycling collectors'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "961 Recycling collectors") 
 
model <-lm (log(salary) ~ sex * year_n + edulevel, data = temp)
 
plot_model(model, type = "pred", terms = c("year_n", "sex"))
```

```{r, echo = TRUE, fig.cap = 'Highest F-value interaction sex, year and education level, Physical and engineering science technicians'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "311 Physical and engineering science technicians") 
 
model <-lm (log(salary) ~ sex * year_n * eduyears, data = temp)
 
plot_model(model, type = "pred", terms = c("year_n", "eduyears", "sex"))	
```

```{r, echo = TRUE, fig.cap = 'Lowest F-value interaction sex, year and education level, Architects and surveyors'} 
temp <- tbnum %>%
  filter(`occuptional  (SSYK 2012)` == "216 Architects and surveyors") 
 
model <-lm (log(salary) ~ sex * year_n * eduyears, data = temp)
 
plot_model(model, type = "pred", terms = c("year_n", "eduyears", "sex"))	
```